Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-07 analysis/analysis.qmd
update references

```{r run-to-update-bibfile, eval=TRUE, include=FALSE}
keys <- rbbt::bbt_detect_citations("analysis.qmd")
citations_only <- grep(pattern="[fig|tbl]-.*", keys, value=TRUE, invert=TRUE)
rbbt::bbt_write_bib(keys=citations_only, 'references.bib', overwrite = TRUE,
                    library_id=rbbt::bbt_library_id('Graphics Research'),
                    translator='bibtex')
```
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      dpi = 72, # we can set the resolution up to 300 later.
                      cache = TRUE) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

@@

# D < C < A < B.
response_67 <- rbind(
    response_6 %>% select(CaseId, labelled, task, responses, weight),
    response_7 %>% select(CaseId, labelled, task, responses, weight)
  ) %>% mutate(
  correct = ifelse(labelled %in% c("AB", "BC", "BD"), responses == "B is bigger", 
              ifelse(labelled %in% c("AC", "AD"), responses == "A is bigger",
                     responses == "C is bigger")),
  correct3 = correct | responses == "They are the same",
  response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct")),
  labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD")),
  signal = ifelse(labelled=="BD", 4, 
             ifelse(labelled=="BC", 3,
               ifelse(labelled %in% c("AB", "AD"), 2, 1)))
)

@@

@fig-response-67 gives an overview of all responses to size assessments of the marked tiles. The panels are ordered according to the difference in signal and hypothesized difficulty from easiest (left) to hardest (right) comparison. We see, that in general the percentage of correct responses decreases from left to right. 
Design does not seem to matter for easy comparisons (BD, BC) or the hard comparison (CD). 
The three comparisons in the middle show interesting patterns, highlighted in the reordered @fig-response-67b.

```{r}
#| label: fig-response-67
#| fig-cap: "Responses for comparisons AB through CD on the three different facetted designs. Responses to the same comparison are shown side-by-side for the three designs. Comparisons are ordered from least difficult (BD) to most difficult (CD). The increasing difficulty of comparisons is reflected in the overall decreasing percentage of correct responses (height of blue tiles). The overlaid rectangles represent 95% confidence intervals."
#| fig-height: 4.25
#| fig-width: 7
#| out-width: 80%

@@

response_67 %>% 
  mutate(
    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
  ) %>% 
  ggplot(aes(x = task)) + 
  geom_bar(aes(fill = response3, weight = weight), position="fill") +
  facet_grid(.~labelled) + 
  theme(legend.position="bottom") + 
  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
  xlab("") + 
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = (response3Correct - 1.96* se.response3Correct),
      xmax = as.numeric(task) + 0.45,
      ymax = (response3Correct + 1.96* se.response3Correct),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = (response3Correct ),
      xend = as.numeric(task) + 0.45,
      yend = (response3Correct )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
      xmax = as.numeric(task) + 0.45,
      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = (1- response3Wrong ),
      xend = as.numeric(task) + 0.45,
      yend = (1-response3Wrong )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
```

@@

response_67 %>% 
  mutate(
    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
  ) %>% 
  ggplot(aes(x = labelled)) + 
  geom_bar(aes(fill = response3, weight = weight), position="fill") +
  facet_grid(.~task) + 
  theme(legend.position="bottom") + 
  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
  xlab("") + 
  geom_rect(
    aes(
      xmin = as.numeric(labelled) - 0.45,
      ymin = (response3Correct - 1.96* se.response3Correct),
      xmax = as.numeric(labelled) + 0.45,
      ymax = (response3Correct + 1.96* se.response3Correct),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(labelled) - 0.45,
      y = (response3Correct ),
      xend = as.numeric(labelled) + 0.45,
      yend = (response3Correct )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  geom_rect(
    aes(
      xmin = as.numeric(labelled) - 0.45,
      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
      xmax = as.numeric(labelled) + 0.45,
      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(labelled) - 0.45,
      y = (1- response3Wrong ),
      xend = as.numeric(labelled) + 0.45,
      yend = (1-response3Wrong )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
```

```{r model-67}
# still putzing around with it
logit_67_correct <- svyglm(correct~task+labelled + task, family=quasibinomial(), design=survey_67)

logit_67_cont <- svyglm(correct~task*signal, family=quasibinomial(), design=survey_67)

>>>>

```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
keys <- rbbt::bbt_detect_citations("analysis.qmd")
citations_only <- grep(pattern="[fig|tbl]-.*", keys, value=TRUE, invert=TRUE)
rbbt::bbt_write_bib(keys=citations_only, 'references.bib', overwrite = TRUE,
                    library_id=rbbt::bbt_library_id('Graphics Research'),
                    translator='bibtex')
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      dpi = 72, # we can set the resolution up to 300 later.
                      cache = TRUE) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

@@

# D < C < A < B.
response_67 <- rbind(
    response_6 %>% dplyr::select(CaseId, labelled, task, responses, weight),
    response_7 %>% dplyr::select(CaseId, labelled, task, responses, weight)
  ) %>% mutate(
  correct = ifelse(labelled %in% c("AB", "BC", "BD"), responses == "B is bigger", 
              ifelse(labelled %in% c("AC", "AD"), responses == "A is bigger",
                     responses == "C is bigger")),
  correct3 = correct | responses == "They are the same",
  response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct")),
  labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD")),
  signal = ifelse(labelled=="BD", 4, 
             ifelse(labelled=="BC", 3,
               ifelse(labelled %in% c("AB", "AD"), 2, 1))),
  task_labelled = factor(paste(task, labelled, sep="_"))
)

@@

@fig-response-67 gives an overview of all responses to size assessments of the marked tiles. The panels are ordered according to the difference in signal and hypothesized difficulty from easiest (left) to hardest (right) comparison. We see, that in general the percentage of correct responses decreases from left to right. 
Design does not seem to matter for easy comparisons (BD, BC) or the hard comparison (CD). 
The three comparisons in the middle show interesting patterns, highlighted in the reordered @fig-response-67b. Colored letters below and above the bars encode significances between pairs of bars of the same color: two proportions are significantly different at a 5% significance level, if they do not have a letter in common [@piephoAlgorithmLetterBasedRepresentation2004]. Significances are adjusted for multiple comparisons as implemented in the `multcomp` package [@hothornSimultaneousInferenceGeneral2008].

```{r model-67, warning = FALSE}
# same model as for plot 
logit_67_correct <- svyglm(correct~task_labelled-1, family=quasibinomial(), design=survey_67)

logit_67_correct3 <- svyglm(correct3~task_labelled-1, family=quasibinomial(), design=survey_67)

ci_confint <- svyby(~correct, by=~task_labelled, design=survey_67, svyciprop)
ci_confint3 <- svyby(~correct3, by=~task_labelled, design=survey_67, svyciprop)

coefs_67 <- broom::tidy(logit_67_correct)
coefs_67_relaxed <- broom::tidy(logit_67_correct3)

# create all pairwise comparisons for tasks and questions.
cross_compare <- glht(logit_67_correct, mcp(task_labelled="Tukey"))
cross_compare3 <- glht(logit_67_correct3, mcp(task_labelled="Tukey"))

# get letters for each confidence 
ci_confint$letters <- cld(cross_compare)$mcletters$Letters
ci_confint3$letters <- cld(cross_compare3)$mcletters$Letters

ci_confint <- ci_confint %>% 
separate_wider_delim(
  task_labelled, delim = "_", names =c("task", "labelled")
)  %>% mutate(
  task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
  labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
)

ci_confint3 <- ci_confint3 %>% 
separate_wider_delim(
  task_labelled, delim = "_", names =c("task", "labelled")
)  %>% mutate(
  task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
  labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
)
```

```{r}
#| label: fig-response-67
#| fig-cap: "Responses for comparisons AB through CD on the three different facetted designs. Responses to the same comparison are shown side-by-side for the three designs. Comparisons are ordered from least difficult (BD) to most difficult (CD). The increasing difficulty of comparisons is reflected in the overall decreasing percentage of correct responses (height of blue tiles). The overlaid rectangles represent 95% confidence intervals. The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter."
#| fig-height: 4.25
#| fig-width: 7
#| out-width: 80%

@@

response_67 %>% 
  mutate(
    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
  ) %>% 
  ggplot(aes(x = task)) + 
  geom_bar(aes(fill = response3, weight = weight), position="fill") +
  facet_grid(.~labelled) + 
  theme(legend.position="bottom") + 
  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
  xlab("") + 
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = (response3Correct - 1.96* se.response3Correct),
      xmax = as.numeric(task) + 0.45,
      ymax = (response3Correct + 1.96* se.response3Correct),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = (response3Correct ),
      xend = as.numeric(task) + 0.45,
      yend = (response3Correct )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  geom_rect(
    aes(
      xmin = as.numeric(task) - 0.45,
      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
      xmax = as.numeric(task) + 0.45,
      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(task) - 0.45,
      y = (1- response3Wrong ),
      xend = as.numeric(task) + 0.45,
      yend = (1-response3Wrong )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
```

@@

response_67 %>% 
  mutate(
    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
  ) %>% 
  ggplot(aes(x = labelled)) + 
  geom_bar(aes(fill = response3, weight = weight), position="fill") +
  facet_grid(.~task) + 
  theme(legend.position="bottom") + 
  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
  xlab("") + 
  geom_rect(
    aes(
      xmin = as.numeric(labelled) - 0.45,
      ymin = (response3Correct - 1.96* se.response3Correct),
      xmax = as.numeric(labelled) + 0.45,
      ymax = (response3Correct + 1.96* se.response3Correct),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(labelled) - 0.45,
      y = (response3Correct ),
      xend = as.numeric(labelled) + 0.45,
      yend = (response3Correct )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  geom_rect(
    aes(
      xmin = as.numeric(labelled) - 0.45,
      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
      xmax = as.numeric(labelled) + 0.45,
      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
    ), 
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = props_ci
  ) +
  geom_segment(
    aes(
      x = as.numeric(labelled) - 0.45,
      y = (1- response3Wrong ),
      xend = as.numeric(labelled) + 0.45,
      yend = (1-response3Wrong )
    ), 
    alpha = 1, colour = "black", linewidth = 0.25,
    data = props_ci
  ) +
  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
```

```{r estimates, eval=FALSE}
coefs_67 %>% 
  mutate(
    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
  gt() %>%
  tab_header("Estimates for Model 3.2",
             " accuracy of responses across pairwise evalulations") %>%
  cols_label(
    estimate = "log odds",
    std.error = "SE",
    statistic = "t-statistic",
    p.value = "p-value"
  ) %>%
  fmt_number(
    columns = c(estimate, std.error),
    decimals = 2
  ) %>%
  fmt_number(
    columns = statistic,
    decimals = 1
  ) %>%
  cols_align(
    columns = p.value,
    align = "right"
  )
```
